---
title: "scATAC explore"
output: html_notebook
---

```{r, results="hide"}
library(Seurat)
library(Signac)
library(rlist)
library(dplyr)
library(ggplot2)
library(GenomeInfoDb)
library(EnsDb.Hsapiens.v86)
set.seed(1234) 
```


## Data 

```{r data}
data.GM12078 = readRDS("GM12078_preproc.rds")
data.REH = readRDS("REH_preproc.rds")
```

# REH 

## Plot QC and Filter 

```{r REH plot}
v_plot.REH = VlnPlot(object = data.REH,
  features = c('pct_reads_in_peaks', 'peak_region_fragments', 'TSS.enrichment', 'blacklist_ratio', 'nucleosome_signal'),
  pt.size = 0.1,
  ncol = 5,
  cols = "orange"
)

v_plot.REH
```

```{r REH filter}
data.REH <- subset(
  x = data.REH,
  subset = peak_region_fragments > 3000 &
  peak_region_fragments < 20000 &
  pct_reads_in_peaks > 15 &
  blacklist_ratio < 0.05 &
  nucleosome_signal < 4 &
  TSS.enrichment > 2
)
```

## Normalization 

Normalizes across cells to correct for differences in cellular sequencing depth. Normalizes across peaks to give higher values to more rare peaks. 
Use all data (q0) to identify to variable features. 

```{r REH norm, results="hide"}
data.REH <- RunTFIDF(data.REH)
data.REH <- FindTopFeatures(data.REH, min.cutoff = 'q0')
```

## Linear Dimensionality Reduction 

Singular value decomposition (SVD) using the features (peaks) selected above. This returns a reduced dimension representation of the object (simialrto PCA).

```{r REH dim reduction}
data.REH <- RunSVD(data.REH)
```

## Depth correlation 

Check correlation between dimensions and seq depth. First component highly correlated -> don't use. 

```{r REH corr depth}
DepthCor(data.REH )
```

## Cluster 
RunUMAP: non linear dim reduction (similar to t-SNE) on linear dim reduction
FindNeighbors: constructs shared nearest neighbour graph 
FindClusters: identifys clusters from SNN graph 

```{r REH cluster, results="hide"}
data.REH <- RunUMAP(object = data.REH, reduction = 'lsi', dims = 2:30)
data.REH <- FindNeighbors(object = data.REH, reduction = 'lsi', dims = 2:30)
data.REH <- FindClusters(object = data.REH, verbose = FALSE, algorithm = 3)
```

```{r REH plot clusters}
DimPlot(object = data.REH, 
        label = TRUE, 
        pt.size = 1.2) 
```


# GM12078

## Plot QC and Filter GM12078

```{r GM plot}
v_plot.GM12078 = VlnPlot(object = data.GM12078,
  features = c('pct_reads_in_peaks', 'peak_region_fragments', 'TSS.enrichment', 'blacklist_ratio', 'nucleosome_signal'),
  pt.size = 0.1,
  ncol = 5,
  cols = "orange"
)

v_plot.GM12078
```

```{r GM filter}
data.GM12078 <- subset(
  x = data.GM12078,
  subset = peak_region_fragments > 3000 &
  peak_region_fragments < 20000 &
  pct_reads_in_peaks > 15 &
  blacklist_ratio < 0.05 &
  nucleosome_signal < 4 &
  TSS.enrichment > 2
)
```

## Normalization 

Normalizes across cells to correct for differences in cellular sequencing depth. Normalizes across peaks to give higher values to more rare peaks. 
Use all data (q0) to identify to variable features. 

```{r GM norm, results="hide"}
data.GM12078 <- RunTFIDF(data.GM12078)
data.GM12078 <- FindTopFeatures(data.GM12078, min.cutoff = 'q0')
```

## Linear Dimensionality Reduction 

Singular value decomposition (SVD) using the features (peaks) selected above. This returns a reduced dimension representation of the object (simialrto PCA).

```{r GM dim reduction}
data.GM12078 <- RunSVD(data.GM12078)
```


## Depth correlation 

Check correlation between dimensions and seq depth. First component highly correlated -> don't use. 

```{r GM corr depth}
DepthCor(data.GM12078)
```

## Cluster 
RunUMAP: non linear dim reduction (similar to t-SNE) on linear dim reduction
FindNeighbors: constructs shared nearest neighbour graph 
FindClusters: identifys clusters from SNN graph 

```{r GM cluster, results="hide"}
data.GM12078 <- RunUMAP(object = data.GM12078, reduction = 'lsi', dims = 2:30)
data.GM12078 <- FindNeighbors(object = data.GM12078, reduction = 'lsi', dims = 2:30)
data.GM12078 <- FindClusters(object = data.GM12078, verbose = FALSE, algorithm = 3)
```

```{r GM plot clusters}
DimPlot(object = data.GM12078, 
        label = TRUE, 
        pt.size = 1.2) 
```

