---
title: "scATAC Enrichment and Pathways"
output: html_notebook
---

Cellcycle analysis using seurats built in gene sets. 

Gene set enrichment, gene ontology over representation and KEGG pathway enrichment for each cluster using ClusterProfiler. 

Summary of results: 
- Cluster 4 and 5 show overrespresentation of cells in G1   
- Cluster 2 shows underrepresentation in G1 and some overrepresentation of S   
- Some terms related to cellcycle and immune respons show up in the GSEA for cluster 3 (up)  
- Many terms related to ion homeostasis show up in GSEA for cluster 1 (up) and 2 (down)   
- Many terms related to immune response show up in GSEA for cluster 4 (down)  
- Many terms related to metabolic processes show up in GSEA for cluster 5 (up)   

```{r, results="hide"}
library(Seurat)
library(Signac)
library(clusterProfiler)
library(enrichplot)
library(ggplot2)
library("org.Hs.eg.db")
library(pathview)
library(RColorBrewer)
library(dplyr)
library(tidyr)
```

```{r, results="hide"}
data.filt = readRDS("REH_GM12078_norm_cluster_seurat.rds")
DA.peaks = read.csv("GM12078_DA_peaks_genes.csv")
DA.peaks
```

```{r, echo=FALSE}
Idents(data.filt) = "sample"
GM12878 = subset(x = data.filt, idents = "GM12078") 
GM12878$new_cluster = ""
GM12878$new_cluster[GM12878$seurat_clusters == "0" | GM12878$seurat_clusters == "10"] = "1"
GM12878$new_cluster[GM12878$seurat_clusters == "5" | GM12878$seurat_clusters == "6" | GM12878$seurat_clusters == "2"] = "2"
GM12878$new_cluster[GM12878$seurat_clusters == "9"] = "3"
GM12878$new_cluster[GM12878$seurat_clusters == "8"] = "4"
GM12878$new_cluster[GM12878$seurat_clusters == "11"] = "5"
```

```{r,results="hide", echo=FALSE}
Idents(GM12878) = "new_cluster"
top20 <- DA.peaks %>% 
  group_by(cluster) %>% 
  slice_max(n = 20, order_by = abs(avg_log2FC))

heatmap = DoHeatmap(GM12878, 
          features = top20[order(top20$cluster),]$query_region, 
          slot = "data") +
  scale_fill_gradient2(
  limits = c(-1,5), 
  mid = "#ffeda0", 
  high = "#e31a1c",
  low = "#2c7fb8",
  space = "Lab",
  na.value = "white",
  guide = "colourbar",
  aesthetics = "fill") 

cluster = DimPlot(object = GM12878,
        group.by = "new_cluster",
        label = TRUE,
        pt.size = 1.2)

cluster
heatmap
```

## Cell Cycle Based on Predicted Gene Activity 

```{r, results="hide"}
DefaultAssay(GM12878) = "gene.act"

#Marker genes included in Seurat 
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

GM12878 <- CellCycleScoring(GM12878, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
```
```{r}
DimPlot(object = GM12878,
        label = TRUE,
        pt.size = 1.2)
```

```{r, results="hide"}
count_phase = GM12878@meta.data %>% count(new_cluster, Phase) %>% spread(key = Phase, value = n)
rownames(count_phase) = count_phase[,1]
count_phase = count_phase[,2:4]

chisq <- chisq.test(count_phase)
corrplot::corrplot(chisq$residuals, is.cor = FALSE)
```


```{r}
count_phase = GM12878@meta.data %>% count(new_cluster, Phase)
ggplot(count_phase, aes(x = new_cluster, y = n, fill = Phase)) +
  geom_bar(stat="identity", position = position_dodge(), color="black") +
  scale_fill_brewer(palette="Blues") +
  theme_minimal()

```

## Cluster 1 

```{r, echo=FALSE}
#log2 fold change 
gene.list = DA.peaks[DA.peaks$cluster == "1",]$avg_log2FC
names(gene.list) = DA.peaks[DA.peaks$cluster == "1",]$gene_id
gene.list = gene.list[sort.list(gene.list, decreasing = T)]
```

#### Gene Ontology Overrepresentation Analysis
No enriched terms found. 

```{r, results="hide"}
ego.1 <- enrichGO(gene = names(gene.list),
                keyType = "ENSEMBL",
                OrgDb = org.Hs.eg.db,
                ont = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.01,
                qvalueCutoff  = 0.05,
                readable      = TRUE)
```

#### Gene Set Enrichement Analysis

```{r, results="hide"}
gse.1 <- gseGO(geneList=gene.list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pAdjustMethod = "none", 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = org.Hs.eg.db)
```

```{r, results="hide", echo=FALSE}
gse.1 = setReadable(gse.1, 'org.Hs.eg.db', 'ENSEMBL')
heatplot(gse.1, foldChange = gene.list, showCategory = 20) +
  scale_fill_gradient2(
  limits = c(-2,2), 
  low = "#225ea8", 
  high = "#e31a1c",
  mid = "#ffffcc",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill") 
```

#### KEGG Pathway Enrichment

No enriched pathways found. 

```{r, results="hide"}
gene.list.entrez = bitr(names(gene.list), 
                        fromType="ENSEMBL", 
                        toType="ENTREZID", 
                        OrgDb="org.Hs.eg.db")

kk.1 <- enrichKEGG(gene = gene.list.entrez$ENTREZID, organism = 'hsa')
```

## Cluster 2

```{r, echo=FALSE}
#log2 fold change 
gene.list = DA.peaks[DA.peaks$cluster == "2",]$avg_log2FC
names(gene.list) = DA.peaks[DA.peaks$cluster == "2",]$gene_id
gene.list = gene.list[sort.list(gene.list, decreasing = T)]
```

 
#### Gene Ontology Overrepresentation Analysis 

Few enriched terms found: cytoplasmic side of membrane, cytoplasmic side of plasma membrane, extrinsic component of cytoplasmic side of plasma membrane and cellular response to alcohol 

```{r, results="hide"}
ego.2 <- enrichGO(gene = names(gene.list),
                keyType = "ENSEMBL",
                OrgDb = org.Hs.eg.db,
                ont = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.01,
                qvalueCutoff  = 0.05,
                readable      = TRUE)
```

#### Gene Set Enrichement Analysis 

```{r, results="hide"}
gse.2 <- gseGO(geneList=gene.list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")
```

```{r, results="hide", echo=FALSE}
gse.2 = setReadable(gse.2, 'org.Hs.eg.db', 'ENSEMBL')
heatplot(gse.2, foldChange = gene.list, showCategory = 20) +
  scale_fill_gradient2(
  limits = c(-2,2), 
  low = "#225ea8",
  high = "#e31a1c",
  mid = "#ffffcc",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
)
```

#### KEGG Pathway Enrichment 

No enriched pathways found. 

```{r, results="hide"}
gene.list.entrez = bitr(names(gene.list), 
                        fromType="ENSEMBL", 
                        toType="ENTREZID", 
                        OrgDb="org.Hs.eg.db")

kk.2 <- enrichKEGG(gene = gene.list.entrez$ENTREZID, organism = 'hsa')
```


## Cluster 3

```{r, echo=FALSE}
#log2 fold change 
gene.list = DA.peaks[DA.peaks$cluster == "3",]$avg_log2FC
names(gene.list) = DA.peaks[DA.peaks$cluster == "3",]$gene_id
gene.list = gene.list[sort.list(gene.list, decreasing = T)]
```

#### Gene Ontology Overrepresentation Analysis 

No enriched terms found. 

```{r, results="hide"}
ego.3 <- enrichGO(gene = names(gene.list),
                keyType = "ENSEMBL",
                OrgDb = org.Hs.eg.db,
                ont = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.01,
                qvalueCutoff = 0.05,
                readable = TRUE)
```

#### Gene Set Enrichement Analysis 

```{r, results="hide"}
gse.3 <- gseGO(geneList=gene.list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")
```

```{r, results="hide", echo=FALSE}
gse.3 = setReadable(gse.3, 'org.Hs.eg.db', 'ENSEMBL')
heatplot(gse.3, foldChange = gene.list, showCategory = 20) +
  scale_fill_gradient2(
  limits = c(-4,4), 
  low = "#225ea8",
  high = "#e31a1c",
  mid = "#ffffcc",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
)
```

#### KEGG pathway enrichment

```{r, results="hide"}
gene.list.entrez = bitr(names(gene.list), 
                        fromType="ENSEMBL", 
                        toType="ENTREZID", 
                        OrgDb="org.Hs.eg.db")

kk.3 <- enrichKEGG(gene = gene.list.entrez$ENTREZID, 
                   organism = 'hsa')
```

```{r, echo=FALSE}
kk.3 = setReadable(kk.3, "org.Hs.eg.db", "ENTREZID")
heatplot(kk.3, showCategory = 20) 
```

## Cluster 4

```{r, echo=FALSE}
#log2 fold change 
gene.list = DA.peaks[DA.peaks$cluster == "4",]$avg_log2FC
names(gene.list) = DA.peaks[DA.peaks$cluster == "4",]$gene_id
gene.list = gene.list[sort.list(gene.list, decreasing = T)]
```

#### Gene Ontology Overrepresentation Analysis 

Only one enriched term found: transcription regulator complex 

```{r, results="hide"}
ego.4 <- enrichGO(gene = names(gene.list),
                keyType = "ENSEMBL",
                OrgDb = org.Hs.eg.db,
                ont = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.01,
                qvalueCutoff = 0.05,
                readable = TRUE)
```

#### Gene Set Enrichement Analysis 

```{r, results="hide"}
gse.4 <- gseGO(geneList=gene.list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")
```

```{r, results="hide", echo=FALSE}
gse.4 = setReadable(gse.4, 'org.Hs.eg.db', 'ENSEMBL')
heatplot(gse.4, foldChange = gene.list, showCategory = 20) +
  scale_fill_gradient2(
  limits = c(-2,2), 
  low = "#225ea8",
  high = "#e31a1c",
  mid = "#ffffcc",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
)
```

#### KEGG Pathway Enrichment Analysis  

```{r, results="hide"}
gene.list.entrez = bitr(names(gene.list), 
                        fromType="ENSEMBL", 
                        toType="ENTREZID", 
                        OrgDb="org.Hs.eg.db")

kk.4 <- enrichKEGG(gene = gene.list.entrez$ENTREZID, organism = 'hsa')
```

```{r}
kk.4 = setReadable(kk.4, "org.Hs.eg.db", "ENTREZID")
heatplot(kk.4)
```

## Cluster 5

```{r, echo=FALSE}
#log2 fold change 
gene.list = DA.peaks[DA.peaks$cluster == "5",]$avg_log2FC
names(gene.list) = DA.peaks[DA.peaks$cluster == "5",]$gene_id
gene.list = gene.list[sort.list(gene.list, decreasing = T)]
```

#### Gene Ontology Overrepresentation Analysis 

No enriched terms found. 

```{r, results="hide"}
ego.5 <- enrichGO(gene = names(gene.list),
                keyType = "ENSEMBL",
                OrgDb = org.Hs.eg.db,
                ont = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.01,
                qvalueCutoff = 0.05,
                readable = TRUE)
```

#### Gene Set Enrichement Analysis 

```{r, results="hide"}
gse.5 <- gseGO(geneList=gene.list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = org.Hs.eg.db, 
             pAdjustMethod = "none")
```

```{r, results="hide", echo=FALSE}
gse.5 = setReadable(gse.5, 'org.Hs.eg.db', 'ENSEMBL')
heatplot(gse.5, foldChange = gene.list, showCategory = 20) +
  scale_fill_gradient2(
  limits = c(-5,5), 
  low = "#225ea8",
  high = "#e31a1c",
  mid = "#ffffcc",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
)
```


#### KEGG Pathway Enrichment

No enriched pathways found. 

```{r, results="hide"}
gene.list.entrez = bitr(names(gene.list), 
                        fromType="ENSEMBL", 
                        toType="ENTREZID", 
                        OrgDb="org.Hs.eg.db")

kk.5 <- enrichKEGG(gene = gene.list.entrez$ENTREZID, organism = 'hsa')
```



